Supplementary Materials

### 2023-05-12 supplementary_materials.Rmd
Update comp table and supplementary materials

    dseq_summary = data.frame(coati = mean(coati),
                      prank = mean(prank),
                      mafft = mean(mafft),
                      clustalo = mean(clustalomega),
                      macse = mean(macse), row.names = "$d_{seq}$")

    ############################################################################
    # one-side wilcoxon rank test
    wtp = wilcox.test(coati, prank)
    wtmf = wilcox.test(coati, mafft)
    wtc = wilcox.test(coati, clustalomega)
    wtmc = wilcox.test(coati, macse)
    wilcoxon = data.frame(coati = NA, prank = wtp$p.value, mafft = wtmf$p.value,
                          clustalo = wtc$p.value, macse = wtmc$p.value,
                          row.names = "Wilcoxon rank test")
    stats_table = rbind(dseq_summary, wilcoxon)

@@

    # fix row and column names and add to stats table
    row.names(alns)[which(row.names(alns) == coati_model)] = "coati"
    colnames(alns) = c("Perfect alignments", "Best alignments", "Imperfect alignments")
    stats_table = rbind(stats_table, t(alns))

@@

    # fix row and column names and add to stats table
    colnames(selection) = c("F1-score pos selection", "F1-score neg selection")
    rownames(selection) = c("coati", "prank", "mafft", "clustalo", "macse")
    stats_table = rbind(stats_table, t(selection))
    colnames(stats_table) = c(coati_model, "PRANK", "MAFFT", "ClustalOmega", "MACSE")

@@

```

```{r echo = FALSE, results = 'asis'}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")

results = "supplementary_data/tri-mg/results/results_summary.csv"
print(kable(summary_table(results, "tri-mg"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7719 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq}$>0 when at least one aligner found a perfect alignment.")))

results = "supplementary_data/tri-ecm/results/results_summary.csv"
print(kable(summary_table(results, "tri-ecm"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-ecm, PRANK, MAFFT, ClustalOmega, and MACSE on 7678 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq}$>0 when at least one aligner found a perfect alignment.")))

results = "supplementary_data/mar-mg/results/results_summary.csv"
print(kable(summary_table(results, "mar-mg"), digits = 5, caption = paste("Accuracy of COATi codon-marginal-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7661 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq}$>0 when at least one aligner found a perfect alignment.")))

results = "supplementary_data/mar-ecm/results/results_summary.csv"
print(kable(summary_table(results, "mar-ecm"), digits = 5, caption = paste("Accuracy of COATi codon-marginal-ecm, PRANK, MAFFT, ClustalOmega, and MACSE on 7710 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq}$>0 when at least one aligner found a perfect alignment.")))
```

@@

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = "\\label{fig:dseq-tri-mg} Comparison of $d_{seq}$ values between COATi codon-triplet-mg and PRANK, MAFFT, ClustalOmega, and MACSE."}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
plot_dseq_main("supplementary_data/tri-mg/results/results_summary.csv", "tri-mg")
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap =  "\\label{fig:dseq-tri-ecm} Comparison of $d_{seq}$ values between COATi codon-triplet-ecm and PRANK, MAFFT, ClustalOmega, and MACSE."}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
plot_dseq_main("supplementary_data/tri-ecm/results/results_summary.csv", "tri-ecm")
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap =  "\\label{fig:dseq-mar-mg} Comparison of $d_{seq}$ values between COATi codon-marginal-mg and PRANK, MAFFT, ClustalOmega, and MACSE."}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
plot_dseq_main("supplementary_data/mar-mg/results/results_summary.csv", "mar-mg")
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap =  "\\label{fig:dseq-mar-ecm} Comparison of $d_{seq}$ values between COATi codon-marginal-ecm and PRANK, MAFFT, ClustalOmega, and MACSE."}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
plot_dseq_main("supplementary_data/mar-ecm/results/results_summary.csv", "mar-ecm")
```

>>>>

    dseq_summary = data.frame(coati = mean(coati, na.rm = TRUE),
                      prank = mean(prank, na.rm = TRUE),
                      mafft = mean(mafft, na.rm = TRUE),
                      clustalo = mean(clustalomega, na.rm = TRUE),
                      macse = mean(macse, na.rm = TRUE),
                      row.names = "$d_{seq}$")

@@

    # fix row and column names and add to stats table
    row.names(alns)[which(row.names(alns) == coati_model)] = "coati"
    colnames(alns) = c("Perfect alignments", "Best alignments", "Imperfect alignments")
    stats_table = rbind(dseq_summary, t(alns))

@@

    # fix row and column names and add to stats table
    colnames(selection) = c("F1-score pos selection", "F1-score neg selection")
    rownames(selection) = c("coati", "prank", "mafft", "clustalo", "macse")
    stats_table = rbind(stats_table, t(selection))
    colnames(stats_table) = c(coati_model, "PRANK*", "MAFFT", "ClustalOmega", "MACSE")
    stats_table = stats_table[, c(coati_model, "MAFFT", "PRANK*", "MACSE", "ClustalOmega")]

@@

wilcoxon_test = function(results_file, coati_model) {
    # read results_summary data
    results = read.csv(results_file)

    # one-side paired wilcoxon rank test
    wtp = wilcox.test(x = coati, y = prank, paired = TRUE, alternative = "less")
    wtmf = wilcox.test(x = coati, y = mafft, paired = TRUE, alternative = "less")
    wtc = wilcox.test(x = coati, y = clustalomega, paired = TRUE, alternative = "less")
    wtmc = wilcox.test(x = coati, y = macse, paired = TRUE, alternative = "less")
    w = c(wtp$p.value, wtmf$p.value, wtc$p.value, wtmc$p.value)
    w
}
```

```{r echo = FALSE, results = 'asis'}
library(kableExtra)
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
results = "supplementary_data/tri-mg/results/results_summary.csv"
k = kable(summary_table(results, "tri-mg"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7719 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq} > 0$ when at least one aligner found a perfect alignment. "), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 50 empty alignments, calculations are based on 7669 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
#
results = "supplementary_data/tri-ecm/results/results_summary.csv"
k = kable(summary_table(results, "tri-ecm"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-ecm, PRANK, MAFFT, ClustalOmega, and MACSE on 7678 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq} > 0$ when at least one aligner found a perfect alignment."), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 41 empty alignments, calculations are based on 7637 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
#
results = "supplementary_data/mar-mg/results/results_summary.csv"
k = kable(summary_table(results, "mar-mg"), digits = 5, caption = paste("Accuracy of COATi codon-marginal-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7661 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq} > 0$ when at least one aligner found a perfect alignment. "), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 50 empty alignments, calculations are based on 7611 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
#
results = "supplementary_data/mar-ecm/results/results_summary.csv"
k = kable(summary_table(results, "mar-ecm"), digits = 5, caption = paste("Accuracy of COATi codon-marginal-ecm, PRANK, MAFFT, ClustalOmega, and MACSE on 7710 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq} > 0$ when at least one aligner found a perfect alignment."), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 35 empty alignments, calculations are based on 7675 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
```

@@

```{r echo = FALSE}
results = "supplementary_data/tri-mg/results/results_summary.csv"
pval_tri_mg = max(wilcoxon_test(results, "tri-mg"))
results = "supplementary_data/tri-ecm/results/results_summary.csv"
pval_tri_ecm = max(wilcoxon_test(results, "tri-ecm"))
results = "supplementary_data/mar-mg/results/results_summary.csv"
pval_mar_mg = max(wilcoxon_test(results, "mar-mg"))
results = "supplementary_data/mar-ecm/results/results_summary.csv"
pval_mar_ecm = max(wilcoxon_test(results, "mar-ecm"))
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-tri-mg} Comparison of log-transformed $d_{seq}$ data with pseudocounts between COATi codon-triplet-mg and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_tri_mg, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/tri-mg/plot_distance.csv", "tri-mg"))
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-tri-ecm} Comparison of log-transformed $d_{seq}$ data with pseudocounts between COATi codon-triplet-ecm and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_tri_ecm, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/tri-ecm/plot_distance.csv", "tri-ecm"))
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-mar-mg} Comparison of log-transformed $d_{seq}$ data with pseudocounts between COATi codon-marginal-mg and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_mar_mg, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/mar-mg/plot_distance.csv", "mar-mg"))
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-mar-ecm} Comparison of log-transformed $d_{seq}$ data with pseudocounts between COATi codon-marginal-ecm and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_mar_ecm, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/mar-ecm/plot_distance.csv", "mar-ecm"))
```
